---
title: "Load Packages and Data"
format: html
previous:
  href: "code-learning_outcomes.qmd"
  text: "Learning Outcomes"
next:
  href: "code-prepare_data.qmd"
  text: "Prepare Data for Edges"
---

You can either paste chunks of code into the RStudio console, or paste it into a R script file and click run to run each line. You can put chunks into parentheses `{}` and it will run the whole chunk in sequence when you click run.

Copy and run the following in RStudio to load your packages:

```{r, eval=FALSE}
library(sna)
library(readr)
library(igraph)
library(tidyverse)
library(tidygraph)
library(visNetwork)
```

To simulate using your own dataset, [download this .csv](ADD%20LINK) to your local directory or build a csv and save it using the sample data linked in the index.

Now you can use the following command which uses `readr` to choose the file from your directory and open it into R.

```{r, eval=FALSE}
# Bring the dataset into R as a dataframe
input_data<-read.csv(file.choose())
```

The contents of your .csv file will now be stored as an object called `input_data` and can be viewed in the "Environment" tab. You should see that there are 30 observations and 8 variables in the shared example.

------------------------------------------------------------------------

::: nav-footer
```{=html}
<a href="{{< meta previous.href >}}" aria-label="Go to previous page: {{< meta previous.text >}}">
  <span aria-hidden="true">«</span> {{< meta previous.text >}}
</a>

<a href="{{< meta next.href >}}" aria-label="Go to next page: {{< meta next.text >}}">
  {{< meta next.text >}} <span aria-hidden="true">»</span>
</a>
```
:::
