---
title: "Create Network Graph"
format: html
previous:
  href: "code-calculate.qmd"
  text: "Calculate Network Measures"
next:
  href: "code-data_vis.qmd"
  text: "Data Visualisation"
---

We can now visualise the network.

There are different ways to plot network graphs depending on what you want to show.

## Basic network graph

The simplest way to view the network is to plot it directly:

```{r, eval=FALSE}
plot(g,
     vertex.size = degree_centrality,
     vertex.label.cex = 0.7,
     edge.arrow.size = 0.2,
     main = "Network graph")
```

This shows the network structure, with node size scaled by degree centrality.

## Improving node labels

Following this approach, node labels will use the index IDs, which can be difficult to interpret. Your dataset includes a column with descriptive labels `name`, so you can replace the IDs. To ensure labels are matched correctly to nodes, we explicitly map `input_data$id` to `input_data$name`.

```{r, eval=FALSE}
# Create lookup table (id → name)
lookup <- input_data %>%
  select(id, name) %>%
  distinct()

# Match graph node names to readable labels
V(g)$label <- lookup$name[match(V(g)$name, lookup$id)]

```

Now replot your network graph.

## Network with communities

We can also detect and visualise communities within the network:

```{r, eval=FALSE}
communities <- cluster_walktrap(g)
plot(communities, g, 
     vertex.size = degree_centrality,
     vertex.label.cex = 0.7,
     edge.arrow.size = 0.2,
     edge.curved = 0.2,
     main = "UoJelly pockets of excellence for open and reproducible practices")
```

This highlights groups of nodes that are more closely connected to each other.

------------------------------------------------------------------------

::: nav-footer
```{=html}
<a href="{{< meta previous.href >}}" aria-label="Go to previous page: {{< meta previous.text >}}">
  <span aria-hidden="true">«</span> {{< meta previous.text >}}
</a>

<a href="{{< meta next.href >}}" aria-label="Go to next page: {{< meta next.text >}}">
  {{< meta next.text >}} <span aria-hidden="true">»</span>
</a>
```
:::
