---
title: "Create a Report"
format: html
previous:
  href: "code-data_vis.qmd"
  text: "Data Visualisation"
next:
  href: "../index.qmd"
  text: "Home"
---

We will now create a report containing your interactive network graph.

## Create a Quarto Document

In RStudio, go to:

**File → New File → Quarto Document**

-   Give your document a title  
-   Add yourself as the author  
-   Keep **HTML** selected  
-   Keep **Knitr** as the engine  

### YAML header

The section at the top of the document, enclosed by dashes `---`, is the YAML header.

This controls the document metadata and output.

You can edit the title like this:

``` {r, eval=FALSE}
---
title: "University of Jelly Network"
format: html
---
```

### Add content

In the body of the document, you can add headings and text.

Headings are created using hashtags:

-   `#` Main title  
-   `##` Section  
-   `###` Subsection  

For example:

``` {r}
## Summary
```

Add 2–3 sentences describing your network.


### Add your network graph

To insert your graph, create a code chunk:

```` {r, eval=FALSE}
```{r, echo=FALSE}
# <YOUR CODE HERE>
```
````

This will display the graph without showing the code.


::: {.callout-note}
## Optional: include code

If you want to include your code in the report:

-   Use `echo=TRUE` to show and run code  
-   Use `eval=FALSE` to show code without running it  
:::


## Render Your Report

Click **Render** in RStudio to generate your report.

::: {.callout-note}
## Output

Your report should include:

- a title  
- a short summary (2–3 sentences)  
- your interactive network graph  

:::

You have now created a reproducible network analysis report.

------------------------------------------------------------------------

::: nav-footer
```{=html}
<a href="{{< meta previous.href >}}" aria-label="Go to previous page: {{< meta previous.text >}}">
  <span aria-hidden="true">«</span> {{< meta previous.text >}}
</a>

<a href="{{< meta next.href >}}" aria-label="Go to next page: {{< meta next.text >}}">
  {{< meta next.text >}} <span aria-hidden="true">»</span>
</a>
```
:::