---
title: "Network Structure"
format: html
previous:
  href: "theory-learning_outcomes.qmd"
  text: "Learning Outcomes"
next:
  href: "theory-degree.qmd"
  text: "Centrality"
---

This section introduces ways of describing the structure of a network.

Understanding structure helps us interpret how nodes are connected and how information or influence might flow.

### Structural holes

A structural hole exists when two nodes (or groups) are not directly connected.

A node that connects these otherwise unconnected parts of the network can act as a broker.

-   brokers may control or mediate information\
-   they can connect groups that would otherwise remain separate

```{r, echo=FALSE, fig.height=3}
library(igraph)

set.seed(42)

# Create two clusters
cluster1 <- make_full_graph(4)
cluster2 <- make_full_graph(4)

# Combine clusters
g <- disjoint_union(cluster1, cluster2)

# Add broker node (node 9)
g <- add_vertices(g, 1)

# Connect broker
g <- add_edges(g, c(4, 9, 9, 5))

# Layout
layout <- layout_with_fr(g)
layout[,2] <- layout[,2] * 0.5   # compress vertical space

# Remove built-in labels
V(g)$label <- ""

# Colours
V(g)$color <- "grey"
V(g)$color[1:4] <- "lightblue"
V(g)$color[5:8] <- "lightgreen"
V(g)$color[9]   <- "tomato"

# Plot
par(mar = c(1.5, 1.5, 1.5, 1.5))

plot(g,
     layout = layout,
     vertex.size = 8,
     vertex.label = V(g)$label,
     edge.width = 2,
     asp = 0,
     main = "Structural Hole with Broker")

# ---- LABEL PLACEMENT ----


bx <- layout[9, 1]
by <- layout[9, 2]

# Graph centre
center <- colMeans(layout)

# Direction: from centre → broker
dx <- bx - center[1]
dy <- by - center[2]

# Normalize
len <- sqrt(dx^2 + dy^2)
dx <- dx / len
dy <- dy / len

# Scale relative to layout spread
xrange <- diff(range(layout[,1]))
yrange <- diff(range(layout[,2]))

offset_scale <- 0.15 * min(xrange, yrange)

label_x <- bx + dx * offset_scale
label_y <- by + dy * offset_scale

text(label_x, label_y,
     labels = "Broker",
     col = "black",
     cex = 1,
     font = 2)

```

::: callout-note
Brokers aren’t necessarily the most connected — they are the most strategically positioned.
:::

These two clusters are not directly connected. The only way information can pass between them is through the broker which acts as a connecting tie between two otherwise disconnected parts of the network. It is possible to have two reciprocal brokers with a bridging tie.

### Bridging ties

Bridging ties are connections that link different parts of a network.

These ties are often:

-   less frequent or weaker\
-   but important for connecting otherwise separate groups

Bridging ties help networks become more connected overall.

```{r, echo=FALSE, fig.height=3}
library(igraph)

set.seed(42)

# Create two clusters
cluster1 <- make_full_graph(4)
cluster2 <- make_full_graph(4)

# Combine clusters
g <- disjoint_union(cluster1, cluster2)

# Add bridging tie
g <- add_edges(g, c(4, 5))

# Layout
layout <- layout_with_fr(g)
layout[,2] <- layout[,2] * 0.5

V(g)$label <- ""

# Colour nodes
V(g)$color <- ifelse(V(g) <= 4, "lightblue", "lightgreen")

# Identify bridging edge
edge_ends <- ends(g, E(g), names = FALSE)

bridge_edge <- which(
  (edge_ends[,1] == 4 & edge_ends[,2] == 5) |
  (edge_ends[,1] == 5 & edge_ends[,2] == 4)
)

# Style edges
E(g)$color <- "grey80"
E(g)$width <- 1
E(g)$color[bridge_edge] <- "red"
E(g)$width[bridge_edge] <- 4

# Plot
par(mar = c(1.5, 1.5, 1.5, 1.5))

plot(g,
     layout = layout,
     vertex.size = 30,
     vertex.label = V(g)$label,
     edge.width = E(g)$width,
     edge.color = E(g)$color,
     main = "Bridging Tie Between Two Clusters")

# ---- LABEL PLACEMENT ----

coords <- layout[c(4, 5), ]

mid_x <- mean(coords[,1])
mid_y <- mean(coords[,2])

dx <- coords[2,1] - coords[1,1]
dy <- coords[2,2] - coords[1,2]

offset_x <- -dy
offset_y <- dx

# Normalize
len <- sqrt(offset_x^2 + offset_y^2)
offset_x <- offset_x / len
offset_y <- offset_y / len

# ---- SCALE-AWARE OFFSET ----
usr <- par("usr")
xrange <- usr[2] - usr[1]
yrange <- usr[4] - usr[3]
offset_scale <- 0.05 * min(xrange, yrange)

label_x <- mid_x + offset_x * offset_scale
label_y <- mid_y + offset_y * offset_scale

text(label_x, label_y,
     labels = "Bridging tie",
     col = "red",
     cex = 1,
     font = 2)
```

### Density

Density describes how connected a network is overall.

It is calculated as:

-   the number of actual connections\

-   divided by the number of possible connections

-   High density → many connections\

-   Low density → fewer connections

Dense networks tend to be more tightly connected.

------------------------------------------------------------------------

::: nav-footer
```{=html}
<a href="{{< meta previous.href >}}" aria-label="Go to previous page: {{< meta previous.text >}}">
  <span aria-hidden="true">«</span> {{< meta previous.text >}}
</a>

<a href="{{< meta next.href >}}" aria-label="Go to next page: {{< meta next.text >}}">
  {{< meta next.text >}} <span aria-hidden="true">»</span>
</a>
```
:::
