---
title: "Symmetric vs. Asymmetric"
format: html
previous:
  href: "intro-dyad.qmd"
  text: "The Dyad"

next:
  href: "intro-dichotomous_vs_valued.qmd"
  text: "Dichotomous vs. Valued"
---

### Symmetric

So far, we have been looking at **symmetric relationships**. These are dyads with ties which have **no direction** – i.e. the relationship is mutual.

In a symmetric dyad:

-   If Node A is connected to Node B,\
-   then Node B is equally connected to Node A.

Symmetric networks are often used when:

-   the relationship is inherently mutual (e.g. collaboration)
-   or when data is simplified into a **binary condition** (i.e. tie vs. no tie).

#### Example: collaboration network

```{r echo=FALSE, fig.height=2}
library(igraph)
library(ggraph)
library(tidygraph)

edges <- data.frame(
  from = "N1",
  to   = "N2"
)

nodes <- data.frame(
  name = c("N1", "N2"),
  label = c("Research Group A", "Research Group B")
)

g <- graph_from_data_frame(edges, vertices = nodes, directed = FALSE)

ggraph(as_tbl_graph(g), layout = "linear") +
  geom_edge_link(width = 1) +
  geom_node_point(size = 6) +
  geom_node_text(aes(label = label), vjust = -1.5, size = 4) +
  theme_void()
```

::: callout-note
## Jargon

Other ways of describing symmetric relationships:

-   mutual dyad\
-   undirected ties\
-   reciprocal relationship (in interpretation, though not always observed directly).
:::

### Asymmetric

**Asymmetric relationships** introduce **direction**.

In an asymmetric dyad, a tie goes *from one node to another*, but not necessarily in return:

-   Node A → Node B **does not imply** Node B → Node A

These are also known as **directed ties**, and are typically shown using arrows.

Asymmetric networks are useful when relationships involve **flow, influence, or action**.

::: callout-note
Examples include:

-   communication (who sends messages to whom)\
-   citation networks (which paper cites another)\
-   customer relationships\
-   supervision or reporting structures
:::

#### Example: directed relationship

```{r echo=FALSE, fig.height=2}


edges <- data.frame(
  from = "N1",
  to   = "N2"
)

nodes <- data.frame(
  name = c("N1", "N2"),
  label = c("Sender", "Receiver")
)

g <- graph_from_data_frame(edges, vertices = nodes, directed = TRUE)

ggraph(as_tbl_graph(g), layout = "linear") +
  geom_edge_link(
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, "mm")
  ) +
  geom_node_point(size = 6) +
  geom_node_text(aes(label = label), vjust = -1.5, size = 4) +
  theme_void()
```

::: callout-note
## Jargon

Other ways of describing asymmetric relationships:

-   directed dyad\
-   non-reciprocal tie\
-   sender → receiver relationship\
:::

------------------------------------------------------------------------

::: nav-footer
```{=html}
<a href="{{< meta previous.href >}}" aria-label="Go to previous page: {{< meta previous.text >}}">
  <span aria-hidden="true">«</span> {{< meta previous.text >}}
</a>

<a href="{{< meta next.href >}}" aria-label="Go to next page: {{< meta next.text >}}">
  {{< meta next.text >}} <span aria-hidden="true">»</span>
</a>
```
:::
